---
title: "CalEnviroScreen describe"
date: "2/23/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(sf)
library(corrplot)
library(Hmisc)
library(splines)
library(mgcv)
library(viridis)
library(gridExtra)
library(knitr)

```


```{r, message=FALSE, warning=FALSE, echo=FALSE}

## Load data


dat<- st_read(dsn = "C:/Users/dnmed/OneDrive - cumc.columbia.edu/AAMEHS/OEHHA", 
                layer = "CES4 Final Shapefile", geometry_column = "geometry", quiet = TRUE)

#replace -999 with NA
dat<- dat %>% 
  mutate_if(is.numeric, ~na_if(., -999))

#summary(dat)

dat.dw<- read_csv("C:/Users/dnmed/OneDrive - cumc.columbia.edu/As Cancer Project/Data/ces20drinkingwaterresultsbycontaminant.csv")

#convert to numeric
dat.dw[,6:31]<- apply(dat.dw[,6:31], 2, as.numeric)


#colnames(dat)
#colnames(dat.dw)

dat_merge<- dat %>% full_join(dat.dw[,c(1,6:31)], by=c("Tract"="Census Tract"))

dat_merge_sf<- dat_merge

#drop geometry for summarizing bc it sometimes produces an error
dat_merge<- dat_merge %>% st_drop_geometry()

dat_merge<- dat_merge %>% rename(Lead_water=Lead.y, Lead_house= Lead.x)

```

# Summarize data 


## Air pollution

CalEnviroScreen 4.0 was released October 2021. The data includes estimates of ozone, PM2.5, diesel PM, toxic releases from industrial emissions and traffic in California at the census-tract level. Definitions are as follows: 

**Ozone:** Mean of summer months (May-October) of the daily maximum 8-hour ozone concentration (ppm), averaged over three years (2017 to 2019)

**PM2.5:** Annual mean concentration of PM2.5 (weighted average of measured monitor concentrations and satellite observations, µg/m3), over three years (2015 to 2017).

**Diesel PM:** Spatial distribution of gridded diesel PM emissions from on-road and non-road sources 2016 (tons/year).

**Toxic releases:** Toxicity-weighted concentrations of modeled chemical releases to air from facility emissions and off-site incineration (averaged over 2017 to 2019 and including releases from Mexican facilities averaged over 2014 to 2016).

**Traffic:** Sum of traffic volumes adjusted by road segment length (vehicle kilometers per hour) divided by total road length (kilometers) within 150 meters of the census tract (traffic volumes estimates for 2017)

Below, we describe the distributions of these exposures, as well as the Spearman's correlation between exposures and socioeconomic and race/ethnicity covariates. Covariates are estimated from the American Community Survey (5-year estimates, 2015-2019) and expressed as a percentile relative to the distribution across CA census tracts. 



```{r,  message=FALSE, warning=FALSE, echo=FALSE, fig.height= 7, fig.width=9}



df_meltAir<- dat_merge %>%
  dplyr::select(Tract, Ozone, PM2_5, DieselPM, Tox_Rel, Traffic) %>% gather("key", "value", -Tract)
  
kable(df_meltAir %>% group_by(key) %>% summarise(N=n(), 
                                           N_NA= sum(is.na(value)), 
                                           Min= min(value, na.rm = TRUE),
                                           Q1= quantile(value, 0.25, na.rm = TRUE), 
                                           Median= quantile(value,0.5, na.rm = TRUE),
                                           Mean= mean(value, na.rm = TRUE),
                                           Q3= quantile(value, 0.75,na.rm = TRUE),
                                           Max= max(value, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 3))






OzHist<- ggplot(dat_merge) + geom_histogram(aes(x=Ozone), fill="red")
PMHist<- ggplot(dat_merge) + geom_histogram(aes(x=PM2_5), fill="blue")
DPMHist<- ggplot(dat_merge) + geom_histogram(aes(x=DieselPM), fill="green")
ToxHist<- ggplot(dat_merge) + geom_histogram(aes(x=Tox_Rel), fill="purple")
TrafHist<- ggplot(dat_merge) + geom_histogram(aes(x=Traffic), fill="orange")


grid.arrange(OzHist, PMHist, DPMHist, ToxHist, TrafHist, ncol=2)



#correlation 

APcovars<- dat_merge %>% dplyr::select(Ozone, PM2_5, DieselPM, Tox_Rel, Traffic, 
                                     EducatP,PovertyP,UnemplP,HousBurdP, Hispanic,AfricanAm,NativeAm)

corAPcovars<- cor(APcovars,use="complete.obs",method = "spearman")
corrplot.mixed(corAPcovars, lower.col = "black", tl.cex = 0.7)



#AP<- dat_merge %>% dplyr::select(Ozone, PM2_5, DieselPM, Tox_Rel, Traffic)
#corAP<- cor(AP,use="complete.obs",method = "spearman")
#orrplot.mixed(corAP, lower.col = "black")


#Fuller distribution

#quantile(dat_merge$Ozone, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$PM2_5, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$DieselPM, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$Tox_Rel, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$Traffic, seq(0,1,by=0.05), na.rm = TRUE)



#Lead in homes also a good distribution 

#summary(dat_merge$Lead.x)
#hist(dat_merge$Lead.x)
#quantile(dat_merge$Lead.x, seq(0,1,by=0.05), na.rm = TRUE)




```




## Drinking water contaminants

Individual drinking water contaminant levels from CalEnviroScreen are available in the earlier 2.0 version. Census tract IDs from 2010 were used in both 2.0 and 4.0 versions, so data have been successfully merged. Concentrations of arsenic, uranium, cadmium and lead for each census tract were estimated as a population-weighted average using measurements taken from 2008 to 2012. Since hypothesized associations between metals in drinking water and cardiovascular disease are thought to arise from chronic exposure (e.g., atherosclerosis due to arsenic), it is reasonable to use drinking water measurements from the years prior to the rates of CVD evaluated in our study (20017-2019). Concentrations are reported as parts per billion (ppb)

```{r,  message=FALSE, warning=FALSE, echo=FALSE, fig.height= 7, fig.width=9}

#summary(dat_merge)

#Metals in drinking water 

df_meltWater<- dat_merge %>%
  dplyr::select(Tract, Arsenic, Uranium, Cadmium, Lead_water) %>% gather("key", "value", -Tract)
  
kable(df_meltWater %>% group_by(key) %>% summarise(N=n(), 
                                           N_NA= sum(is.na(value)), 
                                           Min= min(value, na.rm = TRUE),
                                           Q1= quantile(value, 0.25, na.rm = TRUE), 
                                           Median= quantile(value,0.5, na.rm = TRUE),
                                           Mean= mean(value, na.rm = TRUE),
                                           Q3= quantile(value, 0.75,na.rm = TRUE),
                                           Max= max(value, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 3))





AsHist<- ggplot(dat_merge) + geom_histogram(aes(x=Arsenic), fill="red")
UHist<- ggplot(dat_merge) + geom_histogram(aes(x=Uranium), fill="blue")
CdHist<- ggplot(dat_merge) + geom_histogram(aes(x=Cadmium), fill="green")
PbHist<- ggplot(dat_merge) + geom_histogram(aes(x=Lead_water), fill="purple")


grid.arrange(AsHist, UHist, CdHist, PbHist, ncol=2)

#correlation 

#metals<- dat_merge %>% dplyr::select(Arsenic, Uranium, Cadmium, Lead_water, Nitrate)
#cormetal<- cor(metals,use="complete.obs",method = "spearman")
#corrplot.mixed(cormetal, lower.col = "black")


Watercovars<- dat_merge %>% dplyr::select(Arsenic, Uranium, Cadmium, Lead_water, Nitrate, 
                                     EducatP,PovertyP,UnemplP,HousBurdP, Hispanic,AfricanAm,NativeAm)

corWatercovars<- cor(Watercovars,use="complete.obs",method = "spearman")
corrplot.mixed(corWatercovars, lower.col = "black", tl.cex = 0.7)


#Fuller distribution

#quantile(dat_merge$Arsenic, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$Uranium, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$Cadmium, seq(0,1,by=0.05), na.rm = TRUE)
#quantile(dat_merge$Lead.y, seq(0,1,by=0.05), na.rm = TRUE)



#dat_merge<- dat_merge %>% mutate(ArsenicQ= cut2(Arsenic, g=4), 
#                   UraniumQ=cut2(Uranium, g=4), 
#                   CadmiumQ=cut2(Cadmium,g=4), 
#                   LeadQ= cut2(Lead.y, g=4))

#table(dat_merge$ArsenicQ)
#table(dat_merge$UraniumQ)
#table(dat_merge$CadmiumQ)
#able(dat_merge$LeadQ)



#Nitrate better distribution 

#summary(dat_merge$Nitrate)
#hist(dat_merge$Nitrate)
#quantile(dat_merge$Nitrate, seq(0,1,by=0.05), na.rm = TRUE)




```

## Cardiovascular disease 

The rate of cardiovascular disease is estimated using "Spatially modeled, age-adjusted rate of emergency department (ED) visits for acute myocardial infarction per 10,000 (averaged over 2015-2017)." 

The timeframe of our outcome overlaps with years of estimated air pollution. Since short-term exposure to air pollution has been associated with CVD, the temporarily of the data is suited for our research questions. 

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height= 7, fig.width=9}


kable(dat_merge %>% summarise(N=n(), 
                                           N_NA= sum(is.na(Cardiovas)), 
                                           Min= min(Cardiovas, na.rm = TRUE),
                                           Q1= quantile(Cardiovas, 0.25, na.rm = TRUE), 
                                           Median= quantile(Cardiovas,0.5, na.rm = TRUE),
                                           Mean= mean(Cardiovas, na.rm = TRUE),
                                           Q3= quantile(Cardiovas, 0.75,na.rm = TRUE),
                                           Max= max(Cardiovas, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 3))


ggplot(dat_merge) + geom_histogram(aes(x=Cardiovas), fill="red")




CVDcovars<- dat_merge %>% dplyr::select(Cardiovas, 
                                     EducatP,PovertyP,UnemplP,HousBurdP, Hispanic,AfricanAm,NativeAm)
corCVDcovars<- cor(CVDcovars,use="complete.obs",method = "spearman")
corrplot.mixed(corCVDcovars, lower.col = "black")







```

## Maps

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

#Ozone, PM2_5, DieselPM, Tox_Rel, Traffic
ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Ozone), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")

```


\pagebreak

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

ggplot(dat_merge_sf) +
  geom_sf(aes(fill = PM2_5), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  
ggplot(dat_merge_sf) +
  geom_sf(aes(fill = DieselPM), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  
ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Tox_Rel), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  
ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Traffic), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
#Arsenic, Uranium, Cadmium, Lead_water 

ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Arsenic), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  

ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Uranium), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  
ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Cadmium), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  
ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Lead.y), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  

ggplot(dat_merge_sf) +
  geom_sf(aes(fill = Cardiovas), color = NA)+
  scale_fill_gradientn(colours=rev(magma(6)), trans="sqrt")
  

```



```{r, fig.height = 5, fig.width = 5, echo=FALSE, eval=FALSE, include=FALSE}

## Quick models

### Low birth weight

#+Hispanic+White+AfricanAm+NativeAm+OtherMult (add in if want to adjust for race/eth)


#Water exposures

WaterExp<- c("Arsenic", "Uranium", "Cadmium", "Lead.y", "Nitrate")


WaterMod <- lapply(WaterExp, function(i) {
      # STRING INTERPOLATION WITH sprintf, THEN CONVERTED TO FORMULA OBJECT
      iformula <- as.formula(sprintf("LowBirtWt~s(%s)+EducatP+PovertyP+UnemplP+HousBurdP", i))  
      gam(iformula, data=dat_merge, family = "quasipoisson") 
  
 
})


plot.gam(WaterMod[[1]])
abline(h=0)
plot.gam(WaterMod[[2]])
abline(h=0)
plot.gam(WaterMod[[3]])
abline(h=0)
plot.gam(WaterMod[[4]])
abline(h=0)
plot.gam(WaterMod[[5]])
abline(h=0)

#Air pollution (and housing lead)

APExp<- c("Ozone", "PM2_5", "DieselPM", "Tox_Rel", "Traffic", "Lead.x")


APMod <- lapply(APExp, function(i) {
      # STRING INTERPOLATION WITH sprintf, THEN CONVERTED TO FORMULA OBJECT
      iformula <- as.formula(sprintf("LowBirtWt~s(%s)+EducatP+PovertyP+UnemplP+HousBurdP", i))  
      gam(iformula, data=dat_merge, family = "quasipoisson") 
  
 
})


plot.gam(APMod[[1]])
abline(h=0)
plot.gam(APMod[[2]])
abline(h=0)
plot.gam(APMod[[3]])
abline(h=0)
plot.gam(APMod[[4]])
abline(h=0)
plot.gam(APMod[[5]])
abline(h=0)
plot.gam(APMod[[6]])
abline(h=0)



```




```{r,  echo=FALSE, eval=FALSE, include=FALSE}


### CVD 


#+Hispanic+White+AfricanAm+NativeAm+OtherMult (add in if want to adjust for race/eth)


#Water exposures

WaterExp<- c("Arsenic", "Uranium", "Cadmium", "Lead.y", "Nitrate")


WaterMod <- lapply(WaterExp, function(i) {
      # STRING INTERPOLATION WITH sprintf, THEN CONVERTED TO FORMULA OBJECT
      iformula <- as.formula(sprintf("Cardiovas~s(%s)+EducatP+PovertyP+UnemplP+HousBurdP", i))  
      gam(iformula, data=dat_merge, family = "quasipoisson") 
  
 
})


plot.gam(WaterMod[[1]])
abline(h=0)
plot.gam(WaterMod[[2]])
abline(h=0)
plot.gam(WaterMod[[3]])
abline(h=0)
plot.gam(WaterMod[[4]])
abline(h=0)
plot.gam(WaterMod[[5]])
abline(h=0)

#Air pollution (and housing lead)

APExp<- c("Ozone", "PM2_5", "DieselPM", "Tox_Rel", "Traffic", "Lead.x")


APMod <- lapply(APExp, function(i) {
      # STRING INTERPOLATION WITH sprintf, THEN CONVERTED TO FORMULA OBJECT
      iformula <- as.formula(sprintf("Cardiovas~s(%s)+EducatP+PovertyP+UnemplP+HousBurdP", i))  
      gam(iformula, data=dat_merge,  family = "quasipoisson") 
  
 
})


plot.gam(APMod[[1]])
abline(h=0)
plot.gam(APMod[[2]])
abline(h=0)
plot.gam(APMod[[3]])
abline(h=0)
plot.gam(APMod[[4]])
abline(h=0)
plot.gam(APMod[[5]])
abline(h=0)
plot.gam(APMod[[6]])
abline(h=0)

#extra 


dat_merge %>% group_by(cut2(Arsenic, cuts = c(0,0.5,1,3,5,10))) %>% summarise(n())

dat_merge<- dat_merge %>% mutate(ArsenicQ=cut2(Arsenic, cuts = c(0,0.5,1,3,5,10)))

dat_merge<- dat_merge %>% mutate(CardiovasINT= as.integer(Cardiovas))

dat_merge<- dat_merge %>% mutate(N=rep(10000, nrow(.)))

mod<- gam(Cardiovas~ArsenicQ+s(Ozone)+s(PM2_5)+EducatP+PovertyP+UnemplP+HousBurdP+offset(log(N)), data=dat_merge,  family = "poisson") 

summary(mod)





```


